Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-08-21 Ben_Manuscripts/structure_paper/Draft.tex
pass over entire paper.

%MRS5: could be punchier. I will keep thinking.  I should be the last author.  Chen the names that Joe and Matt wanttu use (presumably with some middle initial?
\title{Understanding the nanoscale structure of hexagonal phase lyotropic 
liquid crystal membranes}
\author{Benjamin J. Coscia}
\author{Richard D. Noble}
\author{Douglas L. Gin}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\alsoaffiliation{Department of Chemistry and Biochemistry, University of Colorado Boulder, Boulder CO 80309, USA}
\author{Joe Yelk}
\author{Matthew A. Glaser}
\affiliation{Department of Physics, University of Colorado Boulder, Boulder CO, 80309, USA}
\author{Xunda Feng}
\affiliation{Department of Chemical and Environmental Engineering, Yale University, New Haven, Connecticut 06511, USA}
\author{Michael R. Shirts}
\email{michael.shirts@colorado.edu}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\begin{document}

  \graphicspath{{./figures/}}

@@

  Nanostructured porous membranes made from the cross-linked hexagonal phase of
  self-assembled lyotropic liquid crystals (LLCs) are a promising material for
  selective separations. In this work, we investigate an experimentally
  characterized LLC membrane using an atomistic molecular model. We use
  simulated X-ray diffraction (XRD) patterns in order to ensure maximum
  consistency between our structures and experiment. We show that pores are
  likely composed of 5 columns of stacked LLC monomers which surround each
  hydrophilic core. Although this system has been reported as dry, we show that
  small amounts of water are necessary to fully reproduce the experimental XRD
  pattern due to asymmetries introduced by hydrogen bonds between monomer head
  groups and water molecules. We explore the composition and structure of the
  nanopores and reveal that there exists a composition gradient rather than a
  abrupt partition between the hydrophilic and hydrophobic regions. A caveat is
  that the time scales of the dynamics are extremely long for this system,
  requiring starting from a range of alternative initial configurations and
  careful examination of the metastable states observed. The clearer picture of
  the nanoscopic structure of these membranes provided in this study will enable
  a better understanding of the mechanisms of small molecule transport within
  these nanopores.
  \end{abstract}

@@

  The H\textsubscript{II} phase pore geometry (Figure~\ref{fig:assembly}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  The H\textsubscript{II} phase forms at room temperature in the presence of
  c.a.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in literature, has been
  referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014}. 
  % BJC6: maybe not a necessary detail
%  The most
%  promising Q\textsubscript{I} phase for membrane applications forms at 70\degree
%  C when monomer is mixed with c.a.~20 wt\% glycerol\cite{carter_glycerol-based_2012}.
  Q\textsubscript{I} phase membranes consist of a tortuous network of three dimensionally 
  interconnected pores that prevent optimal through-plane transport. The densely packed, 
  non-tortuous and uniform sized pores of H\textsubscript{II} phase membranes represent
  the ideal geometry for achieving high solute flux\cite{matyka_tortuosity-porosity_2008}.
  Despite the promise of the H\textsubscript{II} phase, the hexagonally packed
  liquid crystalline domains, formed when monomers self-assemble, are
  isotropically aligned which is detrimental to membrane permeability.

@@

  Recently, researchers have learned how to macroscopically align the hexagonal
  domains which has revived research into H\textsubscript{II} phase LLC membranes.
  Previously, research efforts were focused on the Q\textsubscript{I} phase,
  whose geometry does not require alignment. In 2014, Feng et al.~showed that one
  can align Col\textsubscript{h} hexagonal domains, created by the monomer Na-GA3C11,
  using a magnetic field with subsequent cross-linking to lock the structure in 
  place\cite{feng_scalable_2014}. In 2016, Feng et al. showed that one could obtain
  the same result by confining neat monomer between PDMS or glass substrates since 
  hexagonal mesophases preferentially anchor perpendicular to both surfaces\cite{feng_thin_2016}.
%  the same result using a second technique termed soft confinement\cite{feng_thin_2016}.
  %MRS8: maybe say what it is, not just what it's named? 
  Current experimental efforts are focused on extending the method to the 
  H\textsubscript{II} phase and characterizing the performance of these newly aligned systems.

@@

	\item \textit{R-double}: A weak intensity line, located at exactly half
	the $q_z$ value of R-$\pi$ ($q_z$ = 0.85 \AA$^{-1}$), corresponds to real
	space periodicity of 7.4 \AA. Since this reflection corresponds to double
	the spacing of R-$\pi$ in real space, we will refer to it as R-double. 
	R-double has been interpreted as 2\textsubscript{1} helical ordering of aromatic
	rings along the $z$ axis\cite{feng_scalable_2014}.
        %meaning if one traces the 
	%positions of the aromatic rings with a helical curve, then for each full turn 
	%in the helix, one will encounter two aromatic rings.

@@

	\item \textit{R-spots}: Within R-alkanes, are four spots of higher
	relative intensity.  Accordingly, we name these reflections R-spots. The
	location of all spots is $\sim\;37^{\circ}$ from the $q_z$ axis in their
	respective quadrants. In many liquid crystal systems one can explain the spots
	as the product of alkane chains tilted with respect to the membrane
	plane\cite{govind_simple_2001}.

@@

  The timescale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system. Previous work has shown
  a coarse-grained model self assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}. Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system (see section S-\ref{S-section:self_assembly}). 
  To bypass the slow self-assembly process, we use Python scripts to assemble monomers
  into a structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:build_procedure}).

  % BJC5: This needs to be modified to reflect columnar build procedure
  \begin{figure}[!htb]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{monomer_diagonal.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.42\textwidth}
  \includegraphics[width=\linewidth]{stacked_monomers.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.35\textwidth}
  \includegraphics[width=\linewidth]{initial_pore.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{initial_unit_cell.png}
  \caption{}
  \end{subfigure}
  \caption{(a) We parameterized a single monomer and annealed it to produce a low energy
		configuration. (b) We assembled monomers into columns by stacking them on top of each 
		other. (c) We duplicated each column and rotated them to create hydrophilic pore centers.
		We chose to stack twenty monomers into each column. (d) We duplicated the pores and 
		placed them into a monoclinic unit cell.}\label{fig:build_procedure}
  \end{figure}

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. We find the starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one chooses
  a reasonable value. The pore radius is chosen to be 0.5 nm in our initial
  configurations. The initial distance between pores, within a wide range, also has 
  little effect on the the equilibrated structure. However, one should not start them
  too close or there will be high energy repulsions during early equilibration. We 
  chose an initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental value
  of 4.12 nm. A sensitivity analysis of both parameters is presented in the 
  Supporting information, section S-\ref{S-section:initial_config_dependence}. The 
  distance between vertically stacked monomers, the xy position of monomers with respect 
  to vertically adjacent monomers, and the number of columns per pore do influence the 
  equilibrated structure and require further justification for their choices. We rely on 
  experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration based
  on R-$\pi$ and then allowed the system to readjust during equilibration. We rotated 
  each monomer so the plane of its aromatic head group would be coplanar with the xy plane. We
  explored three different initial monomer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We explore a second system with an initial spacing of 5
  \AA. We briefly explored a third system with an initial spacing of 10
  \AA. However this spacing yields non-physical behavior which is detailed in the 
  Supporting Information, section S-\ref{S-section:initial_dbwl}.

@@

  \subsection{Equilibration}\label{section:equilibration}
  We developed equilibration schemes for creating dry and wet configurations.
  Both schemes start with an initial configuration generated according to the
  previous guidelines. For wet systems, we added the desired concentration of
  water to the initial configuration and carried out equilibration in the same
  way as the dry systems. First, we fixed monomer head groups in place using
  position restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We
  gradually released the position restraints by decreasing the force constants
  over a series of NVT simulations. We allowed the resulting unrestrained
  structure to equilibrate for 5 ns in the NPT ensemble with pressure controlled
  by the berendsen barostat followed by NPT equilibration simulations run for at
  least 400 ns using the Parrinello-Rahman barostat. More equilibration details
  are available in section S-\ref{S-section:equilibration}.

@@

  Using equilibrated structures, we carried out various calculations to
  characterize the system. We defined the point at which a system is equilibrated
  based on when the distance between pores stopped changing.  We determined when
  the distances stopped changing by applying the statistical test,
  \texttt{pymbar.timeseries.detectEquilibration}, to the time series
  \cite{chodera_simple_2016,shirts_statistically_2008}. Typically, the pore-to-pore
  distance equilibrated between 200 and 350 ns. We used data collected after 
  equilibration to do all subsequent analysis.

@@

  We generated simulated XRD patterns based on atomic coordinates in order to
  make a direct experimental comparison. We modeled all atomic coordinates as
  Gaussian spheres of electron density whose maximum and width are defined by
  each atom's atomic number electronic radius respectively. A three dimensional
  Fourier transform (FT) of the array of electron density results in a three
  dimensional structure factor which represents the unit cell in reciprocal
  space. The experimental WAXS measurement was made using a vertically aligned
  film whose pores were oriented perpendicular to the direction of the incident
  X-ray beam. 
  %BJC6: Below is the part I'm having a hard time putting into words. 
  Although the pores are vertically aligned, the crystalline domains are
  still misaligned with respect to the $xy$ plane. To account for this, we averaged
  2D slices of the structure factor at all angles about $|\mathbf{q}| = (0, 0, z)$.

  We normalized all diffraction patterns relative to R-alkanes. We believe that
  the alkane-alkane density, averaged over all angles, is the feature most likely
  to be replicated between experiment and simulation, as atomistic alkane
  parameters are relatively well studied. Other features are dependent on system
  ordering which is likely to have some dependence on initial configuration.  We
  calculated the average intensity within R-alkanes of the experimental pattern,
  $I_{avg}$ and divided all intensities by this value. In this way, the average
  intensity of R-alkanes was equal to 1. When calculating $I_{avg}$, we excluded
  intensities within $\pm$ 30\degree~of the meridional axis defined by $q_r=0$,
  since the simulated patterns differ from experiment in those regions in all
  cases (See Figure~\ref{fig:ralkanes}). Specifically, in contrast to the
  experimental WAXS pattern, R-$\pi$, as it appears in the simulated diffraction
  patterns, intersects with R-alkanes (See Figure~\ref{fig:XRDsim}). We set an
  upper bound on the colorbar by multiplying $I_{avg}$ by a scaling factor, $f$.
  Intensities that appear in the patterns $\geq$ $f\times I_{avg}$ are colored
  uniformly.  We applied the same scaling method to the simulated patterns. We
  carefully chose a scaling factor of $f=3.1$ in order to visibly display all
  features in all patterns.

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{ralkanes.png}
  \caption{R-alkanes}\label{fig:ralkanes}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{rpi_rdouble.png}
  \caption{$q_z|_{r=0}$}\label{fig:rpi_rdouble}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{angular_integration.png}
  \caption{R-spots}\label{fig:rspots}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{saxs_xsection.png}
  \caption{R-spots}\label{fig:saxs_xsection}
  \end{subfigure}
  \caption{(a) We measured the intensity of R-alkanes by calculating the average 
  intensity within the region bounded by $|\mathbf{q}|$ = 1.4~$\AA^{-1}$ and 
  1.57~$\AA^{-1}$ (between 4.0 and 4.5 \AA~in real space). We excluded intensities
  within $\pm$ 30\degree~of the meridional axis defined by $q_r=0$, since the 
  simulated patterns differ from experiment in those regions in all cases. (b) We
  measured the intensity of R-$\pi$ and R-double based on the peaks of the $q_z$
  cross-section of the diffraction pattern. (c) We measured the intensity of R-spots
  by averaging the peak heights produced by radially integrating the patterns within 
  the R-alkanes region. We took the intensity of R-spots as the average of the 
  peak intensities near 37 and 143$\degree$. (d) We measured the intensity of 
  R-pores by measuring the height of the d\textsubscript{100} peak (the leading
  peak closest to $q_r$=0) of the cross-section of the diffraction patterns along
  the $q_r$ axis at $q_z$=0.} \label{fig:xrd_intensities}
  \end{figure}

@@

  Here, $g(z)$ is characterized by an oscillatory function with a period equal to the
  average distance between stacked monomers, and an amplitude that decays exponentially
  (see Figure~\ref{fig:correlation}). The rate of decay is related to the correlation 
  length, L, between monomer head groups. We estimated L by fitting the peaks of $g(z)$
  to a decaying exponential function of the form:
  \begin{equation}
  	Ae^{-z/L}
  	\label{eqn:decaying_exponential}
  \end{equation}
  where A is a fitting parameter for amplitude, $z$ is the independent variable of $g(z)$ 
  and L is the fit correlation length. We calculated the error in the estimated value
  of L as the square root of the diagonal entry of the covariance matrix of 
  optimized fit parameters.  %BJC6: reference scipy curve_fit?

@@

  In order to gain a deeper understanding of discrepancies between the experimental and
  and simulated R-$\pi$ reflection, we used a simplified baseline model where we 
  represent each monomer as a single point scatterer located at the center of mass
  of its head group. In order to make a system approximately similar to our equilibrated 
  simulation geometries, we created 4 hexagonally packed pore regions spaced 42.5 \AA~ 
  apart, each with 5 columns of scatterers spaced 4.4 \AA~apart in the $z$-direction. We
  built the system 4 times taller in the $z$-direction (80 scatterers per column) in order
  to access higher $q_z$ resolution.

@@

  We approximated interactions between particles by correlating the $z$-distance 
  between points within each column of point scatterers. We placed points in each
  column by drawing random samples from a multivariate normal distribution defined
  by the mean positions of an equally spaced column of points. We gave the 
  distribution at each point along the column the same standard deviation, 
  $\sigma_z$ and we defined a covariance matrix such that the covariance, $v$, of
  the distance, $d$, between scatterers decays exponentially from $v$ according
  to the equation $ve^{-d/L}$, where $L$ is the correlation length. Unless noted
  otherwise, we used the experimental correlation length of 9.0 $\AA$.

  We calculated averaged R-$\pi$ profiles by simulating the diffraction pattern
  of trajectories consisting of 1000 independent simple system configurations 
  generated with point scatterer placement based on the quenched order seen in
  atomistic simulations. The quenched disorder is greater in  magnitude than thermal 
  disorder (See Table~\ref{table:quenched_disorder} in section~\ref{section:rpi})
  and inherently includes thermal disorder due to the way we calculate it.
  This method assumes that the timescales for large scale rearrangements of the
  system are much longer than what we can feasibly simulate.

@@

  In order to maximally match the experimental membrane synthesis process,
  we created a cross-linking algorithm that one can apply to equilibrated structures. 
  The primary purpose of cross-linking is to create a mechanically robust membrane.
  For that reason, we are not concerned with replicating the kinetics of the reaction, 
  but instead emphasize understanding how much and in what way cross-linking modulates
  the system's structure.

@@

  With the structural understanding gained by these simulations, it becomes
  possible to evaluate transport of various solutes within the system, and apply the
  knowledge gained from this study in order to suggest improvements to the
  existing system as well as to evaluate new unsynthesized LLC systems.

>>>>

%MRS5: could be punchier. I will keep thinking.  I should be the last author.  Chen the names that Joe and Matt wanttu use (presumably with some middle initial?
\title{Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes}
\author{Benjamin J. Coscia}
\author{Richard D. Noble}
\author{Douglas L. Gin}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\alsoaffiliation{Department of Chemistry and Biochemistry, University of Colorado Boulder, Boulder CO 80309, USA}
\author{Joe Yelk}
\author{Matthew A. Glaser}
\affiliation{Department of Physics, University of Colorado Boulder, Boulder CO, 80309, USA}
\author{Xunda Feng}
\affiliation{Department of Chemical and Environmental Engineering, Yale University, New Haven, Connecticut 06511, USA}
\author{Michael R. Shirts}
\email{michael.shirts@colorado.edu}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\begin{document}
%MRS13: we might end up adjusting the author list a bit; I will ask Matt where he thinks he and Joe should go.  I will talk to Rich when we get this going; not clear that he actually has contributed.  But that's my job. 
  \graphicspath{{./figures/}}

@@

  %MRS13: I tweaked this a bit.  Probably could add another sentence to add some of the other important findings.
  Nanostructured porous membranes made from the cross-linked hexagonal phase of
  self-assembled lyotropic liquid crystals (LLCs) are a promising material for
  selective separations. In this work, we investigate an experimentally
  characterized LLC membrane using atomistic molecular modeling. In particular, we compare
  simulated X-ray diffraction (XRD) patterns with experimental XRD to quantify and understand 
  differences between simulation and experiment.
  We find that pores are
  likely composed of 5 columns of stacked LLC monomers which surround each
  hydrophilic core.  We also find that some of the structure previously attributed to 
  monomer tail tilt is likely instead due to ordered tail packing.  
  Although this system has been reported as dry, we show that
  small amounts of water are necessary to fully reproduce the experimental XRD
  pattern due to asymmetries introduced by hydrogen bonds between monomer head
  groups and water molecules. We explore the composition and structure of the
  nanopores and reveal that there exists a composition gradient rather than a
  abrupt partition between the hydrophilic and hydrophobic regions. A caveat is
  that the time scales of the dynamics are extremely long for this system, resulting in 
  simulated structures that appear too ordered, 
  %MRS13: I left out starting from alternative structures since we didn't actually do it.
  %requiring starting from a range of alternative initial configurations and
  and requiring careful examination of the metastable states observed for any conclusions. The clearer picture of
  the nanoscopic structure of these membranes provided in this study will enable
  a better understanding of the mechanisms of small molecule transport within
  these nanopores.
  \end{abstract}

@@

  The H\textsubscript{II} phase pore geometry (Figure~\ref{fig:assembly}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  The H\textsubscript{II} phase forms at room temperature in the presence of
  c.a.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in literature, has been
  referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014}. 
  % BJC6: maybe not a necessary detail
%  The most
%  promising Q\textsubscript{I} phase for membrane applications forms at 70\degree
%  C when monomer is mixed with c.a.~20 wt\% glycerol\cite{carter_glycerol-based_2012}.
  Q\textsubscript{I} phase membranes consist of a tortuous network of three dimensionally 
  interconnected pores that prevent optimal through-plane transport. In contrast, the densely packed, 
  non-tortuous and uniform sized pores of H\textsubscript{II} phase membranes represent
  the ideal geometry for achieving high solute flux\cite{matyka_tortuosity-porosity_2008}.
  However, the hexagonally packed liquid crystalline domains of the H\textsubscript{II} phase generally 
  form isotropically oriented and therefore mutually unaligned domains which are detrimental to membrane 
  permeability. This domain scale misalignment has inhibited further development of the this technology, and 
  research efforts were focused on the Q\textsubscript{I} phase, whose geometry does not require alignment~\cite{dischinger_papers}.

@@

  However, recently researchers have learned how to macroscopically align the hexagonal
  domains which has revived research into H\textsubscript{II} phase LLC membranes.
  In 2014, Feng et al.~showed that one
  can align Col\textsubscript{h} hexagonal domains, created by the monomer Na-GA3C11,
  using a magnetic field with subsequent cross-linking to lock the structure in 
  place\cite{feng_scalable_2014}. In 2016, Feng et al. showed that one could obtain
  the same result by confining neat monomer between PDMS or glass substrates since 
  hexagonal mesophases preferentially anchor perpendicular to both surfaces\cite{feng_thin_2016}.
%  the same result using a second technique termed soft confinement\cite{feng_thin_2016}.
  %MRS8: maybe say what it is, not just what it's named? 
  Current experimental efforts are focused on extending the method to the 
  H\textsubscript{II} phase and characterizing the performance of these newly aligned systems.

@@

	\item \textit{R-double}: A weak intensity line, located at exactly half
	the $q_z$ value of R-$\pi$ ($q_z$ = 0.85 \AA$^{-1}$), corresponds to real
	space periodicity of 7.4 \AA. Since this reflection corresponds to double
	the spacing of R-$\pi$ in real space, we will refer to it as R-double. 
	R-double has been previously interpreted as 2\textsubscript{1} helical ordering of aromatic
	rings along the $z$ axis\cite{feng_scalable_2014}.
        %meaning if one traces the 
	%positions of the aromatic rings with a helical curve, then for each full turn 
	%in the helix, one will encounter two aromatic rings.

@@

	\item \textit{R-spots}: Within R-alkanes, are four spots of higher
	relative intensity.  Accordingly, we name these reflections R-spots. The
	location of all spots is $\sim\,37^{\circ}$ from the $q_z$ axis in their
	respective quadrants. In many liquid crystal systems one can explain the spots
	as the product of alkane chains tilted with respect to the membrane
	plane\cite{govind_simple_2001}.

@@

  The timescale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system. Previous work has shown
  a coarse-grained model self assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}. Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system (see section S-\ref{S-section:self_assembly}). 
  To bypass the slow self-assembly process, we use Python scripts to assemble monomers
  into a structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:build_procedure}).
  %MRS13: Would be nice to say we put the scripts on GitHub.  Additional details about which scripts can be in supporting.
  These initial structures may result in equilibration into slowly-intercoverting metastable states, an issue we will address later in the paper. 
  % BJC5: This needs to be modified to reflect columnar build procedure
  \begin{figure}[!htb]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{monomer_diagonal.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.42\textwidth}
  \includegraphics[width=\linewidth]{stacked_monomers.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.35\textwidth}
  \includegraphics[width=\linewidth]{initial_pore.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{initial_unit_cell.png}
  \caption{}
  \end{subfigure}
  \caption{(a) We parameterized a single monomer and annealed it to produce a low energy
		configuration. (b) We assembled monomers into columns by stacking them on top of each 
		other. (c) We duplicated each column and rotated them to create hydrophilic pore centers.
		We chose to stack twenty monomers into each column. (d) We duplicated the pores and 
		placed them into a monoclinic unit cell with hexagonal symmetry.}\label{fig:build_procedure}
  \end{figure}

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. We find the starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one chooses
  a reasonable value. The pore radius is chosen to be 0.5 nm in our initial
  configurations. The initial distance between pores, within a wide range, also has 
  little effect on the the equilibrated structure. However, one should not start them
  too close or there will be high energy repulsions during early equilibration. We 
  chose an initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental value
  of 4.12 nm. A sensitivity analysis of both parameters is presented in the 
  %MRS13: be consistent on capitalization of ``Supporting'' ``Information'' and ``section''. Look at JPCB for guidance.
  Supporting information, section S-\ref{S-section:initial_config_dependence}. The 
  distance between vertically stacked monomers, the xy position of monomers with respect 
  to vertically adjacent monomers, and the number of columns per pore do influence the 
  equilibrated structure and require further justification for their choices. We rely on 
  experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration based
  on R-$\pi$ and then allowed the system to readjust during equilibration. We rotated 
  each monomer so the plane of its aromatic head group would be coplanar with the xy plane. We
  explored three different initial monomer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We also extensively explored a second system with an initial spacing of 5
  \AA. We briefly explored a third system with an initial spacing of 10
  \AA. However this largest spacing yields non-physical behavior which is detailed in the 
  Supporting Information, section S-\ref{S-section:initial_dbwl}.

@@

  \subsection{Equilibration}\label{section:equilibration}
  We developed equilibration schemes for creating dry and wet configurations.
  Both schemes start with an initial configuration generated according to the
  previous guidelines. For wet systems, we added the desired concentration of
  water to the initial configuration and carried out equilibration in the same
  way as the dry systems. First, we fixed monomer head groups in place using
  position restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We
  gradually released the position restraints by decreasing the force constants
  over a series of NVT simulations. We allowed the resulting unrestrained
  structure to equilibrate for 5 ns in the NPT ensemble with pressure controlled
  by the Berendsen barostat followed by NPT equilibration simulations run for at
  least 400 ns using the Parrinello-Rahman barostat. More equilibration details
  are given in section S-\ref{S-section:equilibration}.

@@

  Using equilibrated structures, we carried out various calculations to
  characterize the system. We defined the point at which a system is equilibrated
  based on when the distance between pores stopped changing.  We determined when
  the distances stopped changing by applying the statistical test of Chodera~\cite{chodera_simple_2016},
  implemented as \texttt{pymbar.timeseries.detectEquilibration}, to the time series. Typically, the pore-to-pore
  distance equilibrated between 200 and 350 ns. We used data collected after 
  equilibration to do all subsequent analysis.

@@

  %MRS13: I will talk to Matt about getting the code in GitHub.
  We generated simulated XRD patterns based on atomic coordinates in order to
  make a direct experimental comparison. We modeled all atomic coordinates as
  Gaussian spheres of electron density whose maximum and width are defined by
  each atom's atomic number electronic radius respectively. A three dimensional
  Fourier transform (FT) of the array of electron density results in a three
  dimensional structure factor which represents the unit cell in reciprocal
  space. The experimental WAXS measurement was made using a vertically aligned
  film whose pores were oriented perpendicular to the direction of the incident
  X-ray beam. 
  %BJC6: Below is the part I'm having a hard time putting into words. 
  Although the pores are vertically aligned, the crystalline domains are
  still misaligned with respect to the $xy$ plane. To account for this, we averaged
  2D slices of the structure factor at all angles about $|\mathbf{q}| = (0, 0, z)$.

  We normalized all diffraction patterns relative to R-alkanes. We believe that
  the alkane-alkane density, averaged over all angles, is the feature most likely
  to be replicated between experiment and simulation, as atomistic alkane
  parameters are relatively well studied. Other features are dependent on system
  ordering which is likely to have some dependence on initial configuration.  We
  calculated the average intensity within R-alkanes of the experimental pattern,
  $I_{avg}$ and divided all intensities by this value. In this way, the average
  intensity of R-alkanes was equal to 1. When calculating $I_{avg}$, we excluded
  intensities within $\pm$ 30\degree~of the meridional axis defined by $q_r=0$,
  since the simulated patterns differ from experiment in those regions in all
  cases (See Figure~\ref{fig:ralkanes}). Specifically, in contrast to the
  experimental WAXS pattern, R-$\pi$, as it appears in the simulated diffraction
  patterns, intersects with R-alkanes (See Figure~\ref{fig:XRDsim}). We set an
  upper bound on the colorbar by multiplying $I_{avg}$ by a scaling factor, $f$.
  Intensities that appear in the patterns $\geq$ $f\times I_{avg}$ are colored
  uniformly.  We applied the same scaling method to the simulated patterns. We
  chose a scaling factor of $f=3.1$ in order to visibly display all
  features in all patterns.

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{ralkanes.png}
  \caption{R-alkanes}\label{fig:ralkanes}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{rpi_rdouble.png}
  \caption{$q_z|_{r=0}$}\label{fig:rpi_rdouble}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{angular_integration.png}
  \caption{R-spots}\label{fig:rspots}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{saxs_xsection.png}
  \caption{R-spots}\label{fig:saxs_xsection}
  \end{subfigure}
  \caption{(a) We measured the intensity of R-alkanes by calculating the average 
  intensity within the region bounded by $|\mathbf{q}|$ = 1.4~$\AA^{-1}$ and 
  1.57~$\AA^{-1}$ (between 4.0 and 4.5 \AA~in real space). We excluded intensities
  within $\pm$ 30\degree~of the meridional axis defined by $q_r=0$, since the 
  simulated spectra overlap with R-$\pi$ in that region. in those regions in all cases. (b) We
  measured the intensity of R-$\pi$ and R-double based on the peaks of the $q_z$
  cross-section of the diffraction pattern. (c) We measured the intensity of R-spots
  by averaging the peak heights produced by radially integrating the patterns within 
  the R-alkanes region. We took the intensity of R-spots as the average of the 
  peak intensities near 37 and 143$\degree$. (d) We measured the intensity of 
  R-pores by measuring the height of the d\textsubscript{100} peak (the leading
  peak closest to $q_r$=0) of the cross-section of the diffraction patterns along
  the $q_r$ axis at $q_z$=0.} \label{fig:xrd_intensities}
  %MRS13: it's confusing to have the same label (R-spots) beneath c and d.  Should be different labels if different data.
  \end{figure}

@@

  Here, $g(z)$ is characterized by an oscillatory function with a period equal to the
  average distance between stacked monomers, and an amplitude that decays exponentially
  (see Figure~\ref{fig:correlation}). The rate of decay is related to the correlation 
  length, L, between monomer head groups. We estimated L by fitting the peaks of $g(z)$
  %MRS13: I assume a linear least squares fit?  Be precise so it's reproducible.
  to a decaying exponential function of the form:
  \begin{equation}
  	Ae^{-z/L}
  	\label{eqn:decaying_exponential}
  \end{equation}
  where A is a fitting parameter for amplitude, $z$ is the independent variable of $g(z)$ 
  and L is the fit correlation length. We calculated the error in the estimated value
  of L as the square root of the diagonal entry of the covariance matrix of 
  optimized fit parameters.  %BJC6: reference scipy curve_fit?

@@

  In order to gain a deeper understanding of discrepancies between the experimental and
  and simulated R-$\pi$ reflection, we used a simplified model where we 
  represent each monomer as a single point scatterer located at the center of mass
  of its head group. In order to make a system approximately similar to our equilibrated 
  simulation geometries, we created 4 hexagonally packed pore regions spaced 42.5 \AA~ 
  apart, each with 5 columns of scatterers spaced 4.4 \AA~apart in the $z$-direction. We
  built the system 4 times taller in the $z$-direction (80 scatterers per column) in order
  to access higher $q_z$ resolution.

@@

  We approximated interactions between particles by correlating the $z$-distance 
  between points within each column of point scatterers. We placed points in each
  column by drawing random samples from a multivariate normal distribution defined
  by the mean positions of an equally spaced column of points. We gave the 
  distribution at each point along the column the same standard deviation, 
  $\sigma_z$ and we defined a covariance matrix such that the covariance, $v$, of
  the distance, $d$, between scatterers decays exponentially from $v$ according
  to the equation $ve^{-d/L}$, where $L$ is the correlation length. Unless noted
  otherwise, we used the experimental correlation length of 9.0 \AA.

  We calculated averaged R-$\pi$ profiles by simulating the diffraction pattern
  of trajectories consisting of 1000 independent simple system configurations 
  generated with point scatterer placement based on the quenched order seen in
  atomistic simulations. The quenched disorder is greater in  magnitude than thermal 
  disorder (See Table~\ref{table:quenched_disorder} in section~\ref{section:rpi}).
  %MRS13: you just said this already, commenting out.
  %and inherently includes thermal disorder due to the way we calculate it.
  This method assumes that the timescales for large scale rearrangements of the
  system are much longer than what we can feasibly simulate.

@@

  In order to better match the experimental membrane synthesis process,
  we created a cross-linking algorithm that one can apply to equilibrated structures. 
  The primary purpose of cross-linking is to create a mechanically robust membrane.
  For that reason, we are not concerned with replicating the kinetics of the reaction, 
  but instead emphasize understanding how much and in what way cross-linking modulates
  the system's structure.

@@

  With the structural understanding gained by these simulations, it will becomes
  possible to evaluate transport of various solutes within the system, and apply the
  knowledge gained from this study in order to suggest improvements to the
  existing system as well as to evaluate new unsynthesized LLC systems.
